--
title: "strawberries"
author: "Yu"
date: "2020/10/14"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(tidyverse)
library(magrittr)
library(kableExtra)

opts_chunk$set(echo = FALSE, 
               warning = FALSE,
               message = FALSE)
```

## Raspberries

```{r}
sberry <- ag_data %>% filter((Commodity=="STRAWBERRIES") & (Period=="YEAR"))
sberry %<>% select(-c(Period, Commodity))   

#### Does every Data Item begin with "
sum(str_detect(sberry$`Data Item`, "^STRAWBERRIES, ")) == length(sberry$`Data Item`)

sberry$`Data Item` <- str_replace(sberry$`Data Item`,"^STRAWBERRIES","STRAWBERRIES#")
sberry$`Data Item` <- str_replace(sberry$`Data Item`,"MEASURED IN","#MEASURED IN")
sberry %<>% separate(`Data Item`, c("B","type", "meas"), sep = "#") 
sberry %<>% separate(type, c("type", "Production"), sep = " - ")
sberry %<>% separate(type, c("d1", "Type"), sep = ",")
sberry %<>% separate(Production, c("Production", "d2"), sep = ",")
sberry %<>% separate(meas, c("Measures", "Avg"), sep = ", ")

# di <- sberry$`Data Item`
# di_m <- str_split(di, ",", simplify=TRUE)
# dim(di_m)
# 
# unique(di_m[,1])
# di_m <- di_m[,2:4]


sberry %<>% select(-c(B,d1))
sberry %<>% select(-d2)

# head(sberry$type, n=20)
# ty <- str_split(sberry$type, " ", simplify=TRUE)
# head(ty, n=20)

sberry[is.na(sberry)] <- " "  ## OK now Data Item has been split into parts

## onto Domain

```


```{r}
# sberry$Domain %>% unique()

sberry %<>% separate(Domain, c("D_left", "D_right"), sep = ", ")

# sberry$D_left %>% unique()
# sberry$D_right %>% unique()

sberry[is.na(sberry)] <- " "

## And now Domain Category


## sberry$`Domain Category` %>% unique()

sberry %<>% separate(`Domain Category`, c("DC_left", "DC_right"), sep = ", ")

## looks like DC_left combines labels

head(sberry$DC_left %>% unique(),n=20)
head(sberry$DC_right %>% unique(), n=20)


## work on DC_left first

sberry %<>% separate(DC_left, c("DC_left_l", "DC_left_r"), sep = ": ")

## sberry$DC_left_l %>% unique()
## sberry$DC_left_r %>% unique()

## now work on DC_right

head(sberry$DC_right %>% unique(), n=20)

sberry %<>% separate(DC_right, c("DC_right_l", "DC_right_r"), sep = ": ") 


sberry[is.na(sberry)] <- " "

##  OK now we need to eliminate the redundancy
```

```{r}

## fine and remove redundant columns
 
## paste(sberry$D_left,sberry$DC_left_l) %>% unique
## returns -- "TOTAL NOT SPECIFIED"   "CHEMICAL CHEMICAL"     "FERTILIZER FERTILIZER"

## remove column sberry$DC_left_l


sberry %<>%  select(-DC_left_l) 

## test

# sum(sberry$D_right == sberry$DC_right_l)
# [1] 7419
# > sberry$DC_left_r %>% unique()
# [1] " "           "(NITROGEN)"  "(PHOSPHATE)" "(POTASH)"    "(SULFUR)"  

## remove column DC_right_l

sberry %<>% select(-DC_right_l)


## Test for lab1, lab2

# paste(sberry$lab1, sberry$lab2) %>% unique()
# [1] "ACRES HARVESTED" "PRODUCTION  "    "YIELD  "         "   "             "APPLICATIONS  " 
# [6] "TREATED  " 


sberry %<>% mutate(label = paste(lab1,lab2)) 

## test for necisity of "chemical" in col D_left


# paste(sberry$D_left, sberry$D_right) %>% unique()
# [1] "TOTAL  "              "CHEMICAL FUNGICIDE"   "CHEMICAL HERBICIDE"   "CHEMICAL INSECTICIDE"
# [5] "CHEMICAL OTHER"       "FERTILIZER  "   


## remove "Chemical" and joint the columns

sberry %<>% mutate(D_left = "CHEMICAL", D_left = "") 



sberry %<>% mutate(Chemical=paste(D_left, D_right)) 

sberry %<>% select(-c(D_left, D_right)) 


sberry %<>% select(Year, State, Type, Production, Measures,Avg, DC_left_r, DC_right_r, Chemical, Value )

```


```{r}

###  Now the problem is that we have entries in both the "what" and "meas" columns
##  that begin  "MEASURED IN"
##  how many are there

## in the column "what"
cnt_1 <- str_detect(sberry$what, "MEASURED IN")
sum(cnt_1)

## in the column "meas"

cnt_2 <- str_detect(sberry$meas, "MEASURED IN")
sum(cnt_2)

## We want to put them all in the same column
## So, we will separate them from their current column and put them into
## two columns -- then we will test to make sure there aren't any overlaps
## and then merge the two columns

## we're going to use PURRR.  We need a simple function that takes a logical
## variable and a second variable.  It returns the second variable if the logical
## variable is true and returns a blank if it is false


f1 <- function(a,b){
  if(a){
    return(b)
  }else{
      return("")
  }
}
#################################################
##  try it

f1(TRUE,"hi")

f1(!TRUE, "hi")

## now try it with map2()

f1_log <- c(FALSE, TRUE, TRUE)
f1_str <- c("one", "two", "three")

map2(f1_log, f1_str, f1)

#########################################################
## now let's separate the "MEASURED IN" entries in the meas column
## form an index of the entries to be separated out

index_meas <- str_detect(sberry$meas, "MEASURED IN")

## verify the first six values against the dats sberry
head(index_meas)
new <- map2(index_meas, sberry$meas, f1)
new <- unlist(new)
head(new, n=20)
######################################################

sberry %<>% mutate(m_in_1 = unlist(map2(index_meas, sberry$meas, f1))) 

# 
#
# b_f <- sberry
# 
#  
#  
#  l1 <- "MEASURED IN LB"
#  l2 <- "howdy"
#  l3 <- "MEASURED IN LB / ACRE"
# 
#  str_replace(c(l1,l2,l3), "MEASURED IN.*$", "")
# 
# tmp <- str_replace(l4, "MEASURED IN.*$", "")
# 
# b_f %<>%  mutate(meas = str_replace(b_f$meas, "MEASURED IN.*$", "" ))
# 
# Check on whether it worked
#
# cnt_l <- str_detect(b_f$meas, "MEASURED IN")
# sum(cnt_l)
#

sberry %<>% mutate(meas = str_replace(sberry$meas, "MEASURED IN.*$", ""))

## Check
cnt_3 <- str_detect(sberry$meas, "MEASURED IN")
sum(cnt_3)


#########################
## Now we will do the same thing with the 
## "what" column  

### index of cells to be isolated
index_what <- str_detect(sberry$what, "MEASURED IN")
sum(index_what)

### create a column of the isolated cells
sberry %<>% mutate(m_in_2 = unlist(map2(index_what, sberry$what, f1))) 

###  eliminate the isolated cells from the original column
sberry %<>% mutate(what = str_replace(sberry$what, "MEASURED IN.*$", ""))

### test that theere are no more "MEASURED IN" cells in the original column
cnt_what <- str_detect(sberry$what, "MEASURED IN")
sum(cnt_what)

### Check for overlaps


sberry %<>% mutate(units = str_trim(paste(m_in_1, m_in_2))) 

sberry$units %>% unique()


```




```{r}

## now let's clean it up 

sberry$Avg %>% unique()  ## rename Avg

sberry$meas %>% unique()  ## rename marketing

sberry$label %>% unique() ## rename harvest 

sberry$DC_left_r %>% unique() # rename chemical_family

tmp <- sberry$DC_right_r %>% unique() # rename materials --213

tmp <- sberry$Value %>% unique() # values

tmp <- sberry$units %>% unique() # Measures



sberry %<>% rename(Chem_family = DC_left_r, Materials = DC_right_r)

colnames(sberry)

sberry %<>% select(Year, State, type, Marketing, 
                   Measures, Avg, Harvest, Chem_family,
                   Materials, Chemical, Value )

str_trim(paste(sberry$Marketing, sberry$Harvest)) %>% unique

###  these belong in one column

sberry %<>% mutate(production = str_trim(paste(Marketing, Harvest)))

sberry %<>% select(Year, State, type, production, Measures,
                   Avg, Chem_family, Materials, Chemical, Value)


## I missed this one !!

sberry %<>% mutate(Chemical = str_trim(paste(Chem_family, Chemical)))

sberry %<>% select(Year, State, Type, Production, Avg, Measures, Materials, Chemical, Value)

write.csv(sberry,"sberry.csv",row.names = F)
